The Mass Distribution of Stellar-Mass Black Holes

### 2010-06-29 paper.tex
Added table for evidence ratios.

\begin{figure}
  \begin{center}
    \plottwo{plots/dist-parameteric}{plots/dist-non-parameteric}
  \end{center}
  \caption{\label{fig:dists} The median values of the black hole mass
    distribution, $p(m|\theta)$, at various masses implied by the
    posterior $p(\theta|d)$ for the models discussed in Subsections
    \ref{subsec:parameteric-models} and
    \ref{subsec:non-parameteric-models}.  Error bars span the 10\% to
    90\% range.  Note that these ``distributions of distributions''
    are not necessarily normalized, and need not be ``shaped'' like
    the underlying model distributions.}
\end{figure}

@@

There are two broad ways to compare models in the context of the MCMC
analysis of Section \ref{sec:models}.  The first way, which we discuss
in Subsection \ref{subsec:direct-evidence}, is to use the MCMC samples
from the various models to estimate the value of the integral in
Equation \eqref{eq:model-evidence-def}.  The second way is to treat
the model literally as a parameter in a second MCMC analysis.  One
then proposes jumps (see the discussion in Section \ref{sec:models})
not only for the $\theta_i$ within a particular model, but from
parameters $\theta_i$ in model $M_i$ to parameters $\theta_j$ in
$M_j$.  This technique is called a ``reversible-jump MCMC''
\citep{Green1995}; we discuss it in Subsection
\ref{subsec:reversible-jump-mcmc}.

@@

\begin{figure}
  \begin{center}
    \plotone{plots/evidence}
  \end{center}
  \caption{\label{fig:direct-evidence} Direct integration and harmonic
    mean evidence for the various models from Section \ref{sec:models}
    (see Subsection \ref{subsec:direct-evidence}).  In increasing
    order along the $x$-axis, the models are the power-law of
    Subsubsection \ref{subsubsec:power-law} (PL), the decaying
    exponential of Subsubsection \ref{subsubsec:exponential} (E), the
    single Gaussian of Subsection \ref{subsubsec:gaussian} (G), the
    double Gaussian of Subsubsection \ref{subsubsec:gaussian} (TG),
    and the one-, two-, three-, four-, and five-bin histogram models
    of Subsection \ref{subsec:non-parameteric-models} (H1, H2, H3, H4,
    H5, respectively).  The cross symbols give the value of the direct
    integration; the plus symbols give the harmonic mean estimator
    with bootstrap estimates of the 10\% and 90\% bounds on the
    harmonic mean.  It appears that the power law and one-bin
    histogram are the two best models, but the error ranges and
    inconsistency between the two techniques for calculating evidence
    cast some doubt on this conclusion.}
\end{figure}

@@

We have performed a reversible-jump MCMC that jumps between all the
models (both parameteric and non-parameteric) in Section
\ref{sec:models}.  The counts for the number of samples in each model
appear in Figure \ref{fig:rj}.  The most favored model is the power
law from Subsubsection \ref{subsubsec:power-law}, followed by the one-
and two-bin histograms from Subsection
\ref{subsec:non-parameteric-models}.  This confirms the results from
the direct integration and harmonic mean evidence estimators in the
last subsection.  Interestingly, the theoretical curve from
\citet{Fryer2001} places fourth in the ranking of evidence.  However,
while the exponential distribution is a factor of 3.9 times less
likely than the best-fit power law model, it is only 1.2 times less
likely than the two best histogram models.  A choice of model prior
that only slightly favored the exponential model over the
non-parameteric histograms would put it in second place.

@@

\begin{figure}
  \begin{center}
    \plotone{plots/rj}
  \end{center}
  \caption{\label{fig:rj} Counts of samples in the parameter space of
    each model from Section \ref{sec:models} in the reversible-jump
    MCMC of Subsection \ref{subsec:reversible-jump-mcmc}.  The models
    are the power law of Subsubsection \ref{subsubsec:power-law} (PL),
    the exponential of Subsubsection \ref{subsubsec:exponential} (E),
    the Gaussian (G) and two Gaussian (TG) models of Subsubsection
    \ref{subsubsec:gaussian}, and the one-, two-, three-, four-, and
    five-bin histograms of Subsection
    \ref{subsec:non-parameteric-models} (H1, H2, H3, H4, H5,
    respectively).  The number of counts in each model is proportional
    to the evidence for that model; therefore, the most favored model
    is the power-law, followed by the one- and two-bin histogram.  }
\end{figure}

>>>>

\begin{figure}
  \begin{center}
    \plottwo{plots/dist-parameteric}{plots/dist-non-parameteric}
  \end{center}
  \caption{\label{fig:dists} The median values of the black hole mass
    distribution, $p(m|\theta)$, at various masses implied by the
    posterior $p(\theta|d)$ for the models discussed in Sections
    \ref{subsec:parameteric-models} and
    \ref{subsec:non-parameteric-models}.  Error bars span the 10\% to
    90\% range.  Note that these ``distributions of distributions''
    are not necessarily normalized, and need not be ``shaped'' like
    the underlying model distributions.}
\end{figure}

@@

There are two broad ways to compare models in the context of the MCMC
analysis of Section \ref{sec:models}.  The first way, which we discuss
in Section \ref{subsec:direct-evidence}, is to use the MCMC samples
from the various models to estimate the value of the integral in
Equation \eqref{eq:model-evidence-def}.  The second way is to treat
the model literally as a parameter in a second MCMC analysis.  One
then proposes jumps (see the discussion in Section \ref{sec:models})
not only for the $\theta_i$ within a particular model, but from
parameters $\theta_i$ in model $M_i$ to parameters $\theta_j$ in
$M_j$.  This technique is called a ``reversible-jump MCMC''
\citep{Green1995}; we discuss it in Section
\ref{subsec:reversible-jump-mcmc}.

@@

\begin{figure}
  \begin{center}
    \plotone{plots/evidence}
  \end{center}
  \caption{\label{fig:direct-evidence} Direct integration and harmonic
    mean evidence for the various models from Section \ref{sec:models}
    (see Section \ref{subsec:direct-evidence}).  In increasing
    order along the $x$-axis, the models are the power-law of
    Section \ref{subsubsec:power-law} (PL), the decaying
    exponential of Section \ref{subsubsec:exponential} (E), the
    single Gaussian of Section \ref{subsubsec:gaussian} (G), the
    double Gaussian of Section \ref{subsubsec:gaussian} (TG),
    and the one-, two-, three-, four-, and five-bin histogram models
    of Section \ref{subsec:non-parameteric-models} (H1, H2, H3, H4,
    H5, respectively).  The cross symbols give the value of the direct
    integration; the plus symbols give the harmonic mean estimator
    with bootstrap estimates of the 10\% and 90\% bounds on the
    harmonic mean.  It appears that the power law and one-bin
    histogram are the two best models, but the error ranges and
    inconsistency between the two techniques for calculating evidence
    cast some doubt on this conclusion.}
\end{figure}

@@

We have performed a reversible-jump MCMC that jumps between all the
models (both parameteric and non-parameteric) in Section
\ref{sec:models}.  The counts for the number of samples in each model
appear in Figure \ref{fig:rj}; relative probabilities of the models
are given in Table \ref{tab:rj}.  The most favored model is the power
law from Section \ref{subsubsec:power-law}, followed by the one-
and two-bin histograms from Section
\ref{subsec:non-parameteric-models}.  This confirms the results from
the direct integration and harmonic mean evidence estimators in the
last subsection.  Interestingly, the theoretical curve from
\citet{Fryer2001} places fourth in the ranking of evidence.  However,
while the exponential distribution is a factor of 3.9 times less
likely than the best-fit power law model, it is only 1.2 times less
likely than the two best histogram models.  A choice of model prior
that only slightly favored the exponential model over the
non-parameteric histograms would put it in second place.

\begin{table}
  \begin{center}
    \begin{tabular}{|l|r|}
      \hline
      Model & Relative Probability \\
      \hline \hline
      Power Law (Section \ref{subsubsec:power-law}) & 0.441660 \\
      \hline
      Histogram (1 Bin, Section \ref{subsec:non-parameteric-models}) &
      0.143435 \\
      \hline
      Histogram (2 Bin, Section \ref{subsec:non-parameteric-models}) &
      0.135854 \\
      \hline 
      Exponential (Section \ref{subsubsec:exponential}) & 0.112514 \\
      \hline
      Histogram (3 Bin, Section \ref{subsec:non-parameteric-models}) &
      0.075592 \\
      \hline
      Histogram (4 Bin, Section \ref{subsec:non-parameteric-models}) &
      0.038839 \\
      \hline 
      Gaussian (Section \ref{subsubsec:gaussian}) & 0.022999 \\ 
      \hline
      Histogram (5 Bin, Section \ref{subsec:non-parameteric-models}) &
      0.021011 \\
      \hline      
      Two Gaussian (Section \ref{subsubsec:gaussian}) & 0.008096 \\
      \hline
    \end{tabular}
  \end{center}
  \caption{\label{tab:rj} Relative probabilities of the various models
    from Section \ref{sec:models} computed from reversible-jump
    MCMC samples.}
\end{table}

@@

\begin{figure}
  \begin{center}
    \plotone{plots/rj}
  \end{center}
  \caption{\label{fig:rj} Counts of samples in the parameter space of
    each model from Section \ref{sec:models} in the reversible-jump
    MCMC of Section \ref{subsec:reversible-jump-mcmc}.  The models
    are the power law of Section \ref{subsubsec:power-law} (PL),
    the exponential of Section \ref{subsubsec:exponential} (E),
    the Gaussian (G) and two Gaussian (TG) models of Section
    \ref{subsubsec:gaussian}, and the one-, two-, three-, four-, and
    five-bin histograms of Section
    \ref{subsec:non-parameteric-models} (H1, H2, H3, H4, H5,
    respectively).  The number of counts in each model is proportional
    to the evidence for that model; therefore, the most favored model
    is the power-law, followed by the one- and two-bin histogram.  }
\end{figure}
